---
title: "UFC Beards"
output: pdf_document
---
A first Brad-Terry analysis of UFC Beards and wins and losses.

First the libraries that we are using
```{r}
library(BradleyTerry2)
library(dplyr)
library(knitr)
library(ggplot2)
source("functions_for_prepping_bt_analysis.R")
```
Then load the data which comes from Mike.  Note the as.is thing which leaves all the character vectors as strings and does not make them into factors.

```{r}
winner<-read.csv("UFC_winner.csv",as.is=T)
loser<-read.csv("UFC_loser.csv",as.is=T)
predictors<-read.csv("UFC_predictors.csv",as.is=T)
#delete fighters with only one fight which are making BTm go crazy
winner<-winner[c(-18,-157),]
loser<-loser[c(-18,-157),]
predictors<-predictors[c(-6,-162),]
predictors$stance[predictors$stance==""]<-"other"
predictors$stance[predictors$stance=="Open Stance "]<-"other"
predictors$stance[predictors$stance=="Switch "]<-"other"
```
Turns out that th weird BradleyTerry thing needs a very specific format to match things up so we label all the unique id columns "ID" and sort the predictor unique ID column to match the order of the levels of the trial-level files.  This is all now done in an external function.

```{r}
beards<-set_up_btm(predictors,winner,loser)
```

Run the main model


```{r}
model1<-BTm(player1=winner,player2=loser,
            formula = ~ prev + as.factor(facehair) + ht[ID] + reach[ID] +   
            (1|ID), id="ID",data=beards)

summary(model1)
```

See if stance matter

```{r}

model2<-BTm(player1=winner,player2=loser,
            formula = ~ prev + as.factor(facehair)  + ht[ID] + reach[ID] + stance[ID] +
            (1|ID), id="ID",data=beards)

summary(model2)
```


Does this result depend on the type of victory? 

2 or less is a TKO or KO

Greater than 3 is submission or decision or some other outcome.  Here we split the dataset into those two categories and show that facial hair still does not have an effect.

Warning: code gets messy below here.  

```{r}
#other wins including decisions and submissions
w1<-subset(winner,winner$method>=3)
w1$ID<-as.factor(as.character(w1$ID))
l1<-subset(loser,loser$method>=3)
l1$ID<-as.factor(as.character(l1$ID))
p1<-subset(predictors,predictors$ID%in%c(as.character(w1$ID),as.character(l1$ID)))
b.out<-set_up_btm(p1,w1,l1)

model.other.outcomes<-BTm(player1=winner,player2=loser,
            formula = ~  as.factor(facehair)  + reach[ID] + 
            (1|ID), id="ID",data=b.out)
summary(model.other.outcomes)


#TKO's and KO's

w1<-filter(winner,method<=2)
l1<-filter(loser,method<=2)
p1<-subset(predictors,predictors$ID%in%c(as.character(w1$ID),as.character(l1$ID)))
b.out<-set_up_btm(p1,w1,l1)

model3<-BTm(player1=winner,player2=loser,
            formula = ~  as.factor(facehair)  + reach[ID] + 
            (1|ID), id="ID",data=b.out)
summary(model3)
```

some code to make a histogram using a simplified version of the model and a work around because of the bug in the BTabilities function.

```{R}

model.graphics<-BTm(player1=winner,player2=loser,
            formula = ~ prev + facehair  + ht[ID] + reach[ID]  +
            (1|ID), id="ID",data=beards)

summary(model.graphics)
out<-get_bt_abilities(model.graphics,predictors)
predictors$ability<-out$abilities
  all<-rbind(winner,loser)
  all$winning<-c(rep("yes",length(winner$ID)),rep("no",length(winner$ID)))
  all$ID<-as.factor(all$ID)
  winner2<-filter(all,winning=="yes")
  loser2<-filter(all,winning=="no")
beardy<-summarize(group_by(all,name),mean(facehair))
predictors$beardy<-beardy$`mean(facehair)`[match(predictors$name,beardy$name)]
all$winning[all$winning=="yes"]<-"win"
all$winning[all$winning=="no"]<-"loss"
all$finish[all$method>2]<-"other"
all$finish[all$method<=2]<-"knockout"
all$outcome<-paste(all$winning,all$finish,sep="_")
prop.ko<-summarize(group_by(all,name),prop_knocked=sum(outcome=="loss_knockout")/length(finish))

predictors$beardy[predictors$beardy==1]<-"clean_shaven"
predictors$beardy[predictors$beardy==2]<-"bearded"
predictors$beardy[!predictors$beardy%in%c("clean_shaven","bearded")]<-"variable_or_other"
table(predictors$beardy)

#pdf("plot_hist.pdf")
ggplot(predictors,aes(x=ability, fill = beardy)) +
  geom_histogram(binwidth = 0.05)
#dev.off()

predictors$prop.ko<-prop.ko$prop_knocked[match(predictors$name,prop.ko$name)]
#pdf("whatAboutThisOne.pdf")
ggplot(predictors,aes(fill=beardy,x=prop.ko))+geom_histogram(binwidth = 0.02)+xlab("Proportion of fights lost by KO or TKO")+ylab("Number of fighters")
#dev.off()
```





